### 2017-02-21 docs/lc-emulsion.tex
Attempt to clean up lc-droplet stress tensor
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig/trunk@3081 40876441-3496-0410-a613-a2fca35b2ae2

The second contribution is the liquid crystal bulk free energy density
(Equation~\ref{eq-lc-fed-bulk}):
\begin{equation}
f_{\rm lc} =
  {\textstyle\frac{1}{2}}A_0 \left(1-\gamma(\phi)/3\right) Q_{\alpha\beta}^2
- {\textstyle\frac{1}{3}}A_0 \gamma(\phi)
  Q_{\alpha\beta} Q_{\beta\pi} Q_{\pi\alpha}
+ {\textstyle\frac{1}{4}} A_0 \gamma(\phi) (Q_{\alpha\beta})^2,
\end{equation}
where the composition enters parametrically through the control
parameter $\gamma = \gamma(\phi)$. This will be discussion further
below. The distortion terms in the
free energy related to the elastic constants $\kappa_0$ and $\kappa_1$
are unchanged (and are as appear in Equation~\ref{equation-lc-gradient-fe}).

@@

The chemical potential $\mu(\phi,Q_{\alpha\beta})$ resulting from the
combined free energy density is the functional derivative of the
total energy with respect to the composition~$\phi$. In the
case that $\gamma(\phi) \sim \delta\phi$ we generate terms
\begin{equation}
A\phi + B\phi^3 -\kappa \partial_{\gamma}\partial_{\gamma}\phi
- \textstyle{\frac{1}{6}} A_0 \delta Q_{\alpha\beta}^2
- \textstyle{\frac{1}{3}} A_0 \delta
                          Q_{\alpha\beta} Q_{\beta\pi} Q_{\pi\alpha}
+ \textstyle{\frac{1}{4}} A_0 \delta (Q_{\alpha\beta}^2)^2
\end{equation}
from $f_{\rm c}(\phi)$ and $f_{\rm lc}(Q_{\alpha\beta})$.
The interfacial or anchoring contribution to the chemical potential is
\begin{equation} 
-W \left(
 \partial_{\alpha}\phi\partial_{\beta}Q_{\alpha\beta}
+\partial_{\beta}\phi\partial_{\alpha}Q_{\alpha\beta}\right)
- 2W Q_{\alpha\beta} \partial_\alpha\partial_\beta\phi.
\end{equation}

@@

The terms in the molecular field resulting from the bulk free energy
density are:
\begin{equation}
-A_0 (1 - \gamma(\phi)/3) Q_{\alpha\beta}
+A_0 \gamma(\phi)\left[ Q_{\alpha\pi} Q_{\pi\beta}
     - {\textstyle\frac{1}{3}} \delta_{\alpha\beta} Q_{\pi\sigma}^2\right]
-A_0 \gamma (\phi)Q_{\pi\nu}^2 Q_{\alpha\beta} \\ \nonumber
\end{equation}
where the terms related to the elastic contants $\kappa_0$ and $\kappa_1$
have been omitted (cf. Equation~\ref{eq-lc-h-full}). The interfacial
contribution to the molecular field is
\begin{equation}
-W \left( \partial_{\alpha}\phi\partial_{\beta}\phi
- {\textstyle\frac{1}{3}}
  \partial_{\pi}\phi \partial_{\pi}\phi \delta_{\alpha\beta} \right).
\end{equation}
As before, relevant contributions to the molecular field have
been rendered traceless.

@@

The term $P_{\alpha\beta}$ includes the binary fluid and coupling
(between LC and binary fluid) contributions and is somewhat more
complicated than in the bare liquid crystal case:
\begin{eqnarray}
\label{eq-lc-emulsion-pab}
P_{\alpha\beta} =
-\left( \phi\frac{\delta  F}{\delta\phi} - F \right) \delta_{\alpha\beta}
-\frac{\delta F}{\delta \left(\partial_{\beta}\phi\right)}\partial_{\alpha}\phi
-\frac{\delta F}{\delta\left(\partial_{\beta}Q_{\gamma\nu}\right)}\partial_{\alpha}Q_{\gamma\nu}.
\end{eqnarray}
where we have re-introduced the full free energy functional $F$.
This term is discussed in the following section.

@@

\begin{equation}
\partial_{\alpha}\left(\phi\mu - F\right)
+\partial_{\beta}\frac{\delta F}{\delta Q_{\pi\nu,\beta}}
 \partial_{\alpha} Q_{\pi\nu}
+\frac{\delta F}{\delta Q_{\pi\nu,\beta}} \partial_{\alpha}\partial_{\beta}
 Q_{\pi\nu}
+\partial_{\beta} \frac{\delta F}{\delta \left(\partial_{\beta}\phi\right)}\partial_{\alpha}\phi
+\frac{\delta F}{\delta \left(\partial_{\beta}\phi\right)}\partial_{\alpha}\partial_{\beta}\phi.
\nonumber
\end{equation}
With some manipulation this can be recast as 
\begin{equation}
-\left( \frac{\partial F}{\partial Q_{\pi\nu}} 
       - \partial_{\beta}\frac{\delta F}{\delta Q_{\pi\nu,\beta}}
 \right) \partial_{\alpha}Q_{\pi\nu}
- \left(\frac{\partial F}{\partial\phi} 
- \partial_{\beta} \frac{\delta F}{\delta \left(\partial_{\beta}\phi\right)}\right)\partial_{\alpha}\phi + \partial_{\alpha}\left(\mu\phi\right)
\nonumber
\end{equation}
which is a local force on the fluid
\begin{equation}
\label{eq-lc-emulsion-hgradq}
H_{\pi\nu} \partial_{\alpha} Q_{\pi\nu} + \phi \partial_{\alpha}\mu.
\end{equation}

The total force on the fluid locally is the divergence of the stress,
where the total stress is $\Pi_{\alpha\beta}$
given by equation~(\ref{eq-lc-emulsion-stress}).
It is sometimes useful to compute the various contributions to the
force in different ways. The contribution from the symmetric and
antisymmetric stress $\sigma_{\alpha\beta}$ and  $\tau_{\alpha\beta}$
can be calculated numerically as a divergence. The contribution from
$P_{\alpha\beta}$ can be computed numerically as a divergence, or via the
analytical form of the divergence Equation~\ref{eq-lc-emulsion-hgradq}.
The  last option is
at the expense of global conservation of momentum.

>>>>

The second contribution is the liquid crystal bulk free energy density
(Equation~\ref{eq-lc-fed-bulk}):
\begin{eqnarray}
f_{\rm lc} &=&
  {\textstyle\frac{1}{2}}A_0 \left(1-\gamma(\phi)/3\right) Q_{\alpha\beta}^2
- {\textstyle\frac{1}{3}}A_0 \gamma(\phi)
  Q_{\alpha\beta} Q_{\beta\pi} Q_{\pi\alpha}
+ {\textstyle\frac{1}{4}} A_0 \gamma(\phi) (Q_{\alpha\beta})^2\nonumber\\
&+&{\textstyle\frac{1}{2}} \kappa_0 (\partial_\alpha Q_{\alpha\beta})^2
+ {\textstyle\frac{1}{2}} \kappa_1
(\epsilon_{\alpha\mu\nu} \partial_\mu Q_{\nu\beta} + 2q_0 Q_{\alpha\beta})^2,
\label{eq-lc-fed}
\end{eqnarray}
where in the first line of Equation~\ref{eq-lc-fed} the composition enters parametrically through the control
parameter $\gamma = \gamma(\phi)$. This will be discussion further
below. The distortion terms in the
second line of Equation~\ref{eq-lc-fed} related to the elastic constants $\kappa_0$ and $\kappa_1$
are unchanged (and are as appear in Equation~\ref{equation-lc-gradient-fe}).

@@

The chemical potential $\mu(\phi,Q_{\alpha\beta})$ resulting from the
combined free energy density is the functional derivative of the
total energy with respect to the composition~$\phi$,

\begin{equation}
\mu = \mu_{\rm c} + \mu_{\rm lc} + \mu_{\rm anchoring} = \frac{\delta F}{\delta \phi} 
= \frac{\partial F}{\partial \phi} - \partial_\alpha \frac{\partial F}{\partial (\partial_\alpha \phi)}  
\end{equation}

In the
case that $\gamma(\phi) \sim \delta\phi$ we generate terms
\begin{equation}
\mu_{\rm c} + \mu_{\rm lc} = A\phi + B\phi^3 -\kappa \partial_{\gamma}\partial_{\gamma}\phi
- \textstyle{\frac{1}{6}} A_0 \delta Q_{\alpha\beta}^2
- \textstyle{\frac{1}{3}} A_0 \delta
                          Q_{\alpha\beta} Q_{\beta\pi} Q_{\pi\alpha}
+ \textstyle{\frac{1}{4}} A_0 \delta (Q_{\alpha\beta}^2)^2
\end{equation}
from $f_{\rm c}(\phi)$ and $f_{\rm lc}(Q_{\alpha\beta})$.
The interfacial or anchoring contribution to the chemical potential is
\begin{equation} 
\mu_{\rm anchoring}=-W \left(
 \partial_{\alpha}\phi\partial_{\beta}Q_{\alpha\beta}
+\partial_{\beta}\phi\partial_{\alpha}Q_{\alpha\beta}\right)
- 2W Q_{\alpha\beta} \partial_\alpha\partial_\beta\phi.
\end{equation}

@@

The molecular field resulting from the combined free energy is the functional
derivative of the total free energy with respect to the tensor order parameter $Q_{\alpha\beta}$,

\begin{equation}
H_{\alpha\beta} = H_{\alpha\beta,{\rm lc}} + H_{\alpha\beta, {\rm anchoring}}  = \frac{\delta F}{\delta Q_{\alpha\beta}} 
= \frac{\partial F}{\partial Q_{\alpha\beta}} - \partial_\gamma \frac{\partial F}{\partial (\partial_\gamma Q_{\alpha\beta})}  
\end{equation}

The terms in the molecular field resulting from the bulk free energy
density of the liquid crystalline part are:
\begin{equation}
H_{\alpha\beta,{\rm lc}} = -A_0 (1 - \gamma(\phi)/3) Q_{\alpha\beta}
+A_0 \gamma(\phi)\left[ Q_{\alpha\pi} Q_{\pi\beta}
     - {\textstyle\frac{1}{3}} \delta_{\alpha\beta} Q_{\pi\sigma}^2\right]
-A_0 \gamma (\phi)Q_{\pi\nu}^2 Q_{\alpha\beta} \\ \nonumber
\end{equation}
where the terms related to the elastic contants $\kappa_0$ and $\kappa_1$
have been omitted (cf. Equation~\ref{eq-lc-h-full}). The interfacial
contribution to the molecular field is
\begin{equation}
H_{\alpha\beta, {\rm anchoring}}  = -W \left( \partial_{\alpha}\phi\partial_{\beta}\phi
- {\textstyle\frac{1}{3}}
  \partial_{\pi}\phi \partial_{\pi}\phi \delta_{\alpha\beta} \right).
\end{equation}
As before, relevant contributions to the molecular field have
been rendered traceless.

@@

The term $P_{\alpha\beta}$ includes the binary fluid and coupling
(between LC and binary fluid) contributions and is somewhat more
complicated than in the bare liquid crystal case:
\begin{eqnarray}
\label{eq-lc-emulsion-pab}
P_{\alpha\beta} =
-\left( \phi\frac{\delta  F}{\delta\phi} - F \right) \delta_{\alpha\beta}
-\frac{\delta F}{\delta \left(\partial_{\beta}\phi\right)}\partial_{\alpha}\phi
-\frac{\delta F}{\delta\left(\partial_{\beta}Q_{\gamma\nu}\right)}\partial_{\alpha}Q_{\gamma\nu},
\end{eqnarray}
where we have re-introduced the full free energy functional $F$.
This term is discussed in the following section.

@@

\begin{equation}
\partial_{\alpha}\left(\phi\mu - F\right)
+\partial_{\beta} \left(\frac{\delta F}{\delta \left(\partial_{\beta}\phi\right)}\right)\partial_{\alpha}\phi
+\frac{\delta F}{\delta \left(\partial_{\beta}\phi\right)}\partial_{\beta}\left(\partial_{\alpha}\phi\right)
+\partial_{\beta}\left(\frac{\delta F}{\delta Q_{\pi\nu,\beta}}\right)
 \partial_{\alpha} Q_{\pi\nu}
+\frac{\delta F}{\delta Q_{\pi\nu,\beta}} \partial_{\beta}\left(\partial_{\alpha} Q_{\pi\nu}\right).
\nonumber
\end{equation}
With some manipulation and using $$\frac{\partial F}{\partial (\partial_\alpha\partial_\beta\phi)} = \frac{\partial F}{\partial (\partial_\alpha Q_{\mu\nu,\beta})}=0$$
this can be recast as 
\begin{eqnarray}
-\partial_\beta P_{\alpha\beta} &=& -\left( \frac{\partial F}{\partial Q_{\pi\nu}} 
       - \partial_{\beta}\frac{\partial F}{\partial Q_{\pi\nu,\beta}}
 \right) \partial_{\alpha}Q_{\pi\nu}
- \left(\frac{\partial F}{\partial\phi} 
- \partial_{\beta} \frac{\partial F}{\partial \left(\partial_{\beta}\phi\right)}\right)\partial_{\alpha}\phi + \partial_{\alpha}\left(\mu\phi\right)\nonumber\\
&=& -\left(\frac{\delta F}{\delta Q_{\pi\nu}}\right) \partial_{\alpha}Q_{\pi\nu} - \left(\frac{\delta F}{\delta \phi}\right) \partial_{\alpha} \phi + \partial_{\alpha}\left(\mu\phi\right)\nonumber\\ 
&=& H_{\pi\nu}\; \partial_{\alpha} Q_{\pi\nu} + \phi \;\partial_{\alpha}\mu,
\label{eq-lc-emulsion-hgradq}
\end{eqnarray}
which is a local force on the fluid.
The total force on the fluid locally is the divergence of the stress,
where the total stress is $\Pi_{\alpha\beta}$
given by equation~(\ref{eq-lc-emulsion-stress}).
It is sometimes useful to compute the various contributions to the
force in different ways. The contribution from the symmetric and
antisymmetric stress $\sigma_{\alpha\beta}$ and  $\tau_{\alpha\beta}$
can be calculated numerically as a divergence. The contribution from
$P_{\alpha\beta}$ can be computed numerically as a divergence, or via the
analytical form of the divergence Equation~\ref{eq-lc-emulsion-hgradq}.
The  last option is
at the expense of global conservation of momentum.
